---
title: "Gracilaria"
author: "Lilith Diener"
date: "2025-02-25"
output: html_document
---

## ...

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(sf)
library(ggplot2)
library(rinat)
library(rosm)
library(ggspatial)
```

Project idea: Establishing the distribution of agar-producing red algae in 
relation to existing MPAs

Red algae species: Gelidium pristoides, Gelidium abbottiorum, 
Gelidium pteridifolium, Gracilaria gracilis

```{r, include=FALSE}
# Call the data directly from iNat
gpr <- get_inat_obs(taxon_name = "Gelidium pristoides",
                   bounds = c(-35, 18, -33.5, 18.5),
                   maxresults = 1000)

#View the first few rows of data
head(gpr)

# Call the data directly from iNat
ga <- get_inat_obs(taxon_name = "Gelidium abbottiorum",
                    bounds = c(-35, 18, -33.5, 18.5),
                    maxresults = 1000)

#View the first few rows of data
head(ga)

# Call the data directly from iNat
gpt <- get_inat_obs(taxon_name = "Gelidium pteridifolium",
                   bounds = c(-35, 18, -33.5, 18.5),
                   maxresults = 1000)

#View the first few rows of data
head(gpt)

# Call the data directly from iNat
gg <- get_inat_obs(taxon_name = "Gracilaria gracilis",
                   bounds = c(-35, 18, -33.5, 18.5),
                   maxresults = 1000)

#View the first few rows of data
head(gg)
```

```{r, include=FALSE}  
# Merge the 4 datasets into a single data frame using rbind()
# rbind() works because all four datasets come from iNat and have the 
# same column structure. If column names differed, I would have had to standardise 
# them first.

# Add a new column before merging to keep track of which species each 
# observation belongs to

# Add species column
gpr$species <- "Gelidium pristoides"
ga$species <- "Gelidium abbottiorum"
gpt$species <- "Gelidium pteridifolium"
gg$species <- "Gracilaria gracilis"

# Merge into one dataset
red_algae_data <- rbind(gpr, ga, gpt, gg)

# Check structure
str(red_algae_data)
```

```{r, include=FALSE}                  
#Filter returned observations by a range of column attribute criteria
gpr <- gpr %>% filter(positional_accuracy<46 & 
                      latitude<0 &
                      !is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")

class(gpr)

ga <- ga %>% filter(positional_accuracy<46 & 
                      latitude<0 &
                      !is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")

class(ga)

gpt <- gpt %>% filter(positional_accuracy<46 & 
                      latitude<0 &
                      !is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")

class(gpt)

gg <- gg %>% filter(positional_accuracy<46 & 
                      latitude<0 &
                      !is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")

class(gg)
```

```{r, include=FALSE}  
#Make the dataframe a spatial object of class = "sf"
red_algae_data <- st_as_sf(red_algae_data, coords = c("longitude", "latitude"), 
                           crs = 4326)

class(red_algae_data) # what class is it?

names(red_algae_data) # note the new "geometry" column
```

```{r, include=FALSE}  
#Plot
ggplot() + geom_sf(data=red_algae_data)
```

DATA RANGE VERGRÖßERN?

```{r, include=FALSE}  
# Add basemap to plot
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=red_algae_data)
```

Note that there are quite a few base layer/tile options that can be set with type = "". Try rosm::osm.types() to see them all.

```{r, include=FALSE}  
# Read in Marine Protected Areas data
mpa 

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
